Elastic Scattering in General Relativistic Ray Tracing for Neutrinos

### 2017-07-05 paper.tex
Finish Momentum Definitions appendix

In practice we use two possible fiducial observers to define the momentum
decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
and the fluid, or comoving observer
$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
Here $t$ is coordinate time and $\alpha$ is lapse
in the standard 3+1 decomposition of the metric
\begin{equation}
  \label{eqn:adm_metric}
  \psi_{\mu\gamma} :=
  \left(
  \begin{matrix}
    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
    \beta_j                      & g_{ij}
  \end{matrix}
  \right).
\end{equation}
We have also introduced
the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
and the projection tensor orthogonal to the normal observer
$g^\mu_\lambda=\psi^\mu_\lambda+u^\mu u_\lambda$,
whose spatial part is the 3-metric, $g_{ij}$.

We specify the neutrino direction in the observer's frame
with two spherical polar angles
\begin{align}
  \ell_i &= q\Omega_i, \\
  \Omega_i &:= (\sin a \cos b,\sin a\sin b,\cos a).
\end{align}
with q and s given by
\begin{align}
  q &= W
\end{align}
And we further specify a rotated frame using two new spherical polar angles
\begin{equation}
  \label{eqn:def_direction_primed_2}
  \Omega_{i'} :=  (\sin A \cos B,\sin A\sin B,\cos A).
\end{equation}
The two frames are related by a standard Euler rotation matrix
\begin{equation}
  \label{eqn:rotation_connection}
  \Omega_i=R^{i'}_i(\theta,\phi)\, \ell_{i'}.
\end{equation}

>>>>

In practice we use two possible fiducial observers to define the momentum
decomposition via Eqn.~\ref{eqn:def_momentum}: the Eulerian,
or normal observer $n_\mu=-\alpha \,\partial_\mu t$,
and the fluid, or comoving observer
$u^\mu=Wn^\mu+v_{\rm E}^\mu$.
Here $t$ is coordinate time and $\alpha$ is lapse
in the standard 3+1 decomposition of the metric
\begin{equation}
  \label{eqn:adm_metric}
  \psi_{\mu\gamma} \rightarrow
  \left(
  \begin{matrix}
    -\alpha^2 + \beta^i \beta_i  & \beta_i \\
    \beta_j                      & g_{ij}
  \end{matrix}
  \right).
\end{equation}
We have also introduced
the fluid Lorentz factor $W=-n_\mu u^\mu=\alpha u^t$,
its Eulerian velocity $v_{\rm E}^\mu=g^\mu_\lambda u^\lambda$
(distinct from its coordinate velocity $v^\mu=u^\mu/u^t$),
and the projection tensor orthogonal to the normal observer
$g^\mu_\lambda=\psi^\mu_\lambda+u^\mu u_\lambda$,
whose spatial part is the 3-metric, $g_{ij}$.

We specify the neutrino direction in the observer's frame
with two spherical polar angles
($a$,$b$) with respect to the simulation cartesian coordinates
\begin{align}
  \ell_\alpha &\rightarrow q(s,\Omega_i), \\
  \Omega_i &\rightarrow (\sin a \cos b,\sin a\sin b,\cos a),
\end{align}
or alternatively the two spherical polar angles
($A$,$B$) with respect to rotated coordinates
\begin{equation}
  \label{eqn:def_direction_primed_2}
  \Omega_{i'} \rightarrow  (\sin A \cos B,\sin A\sin B,\cos A).
\end{equation}
The two coordinate systems are related by a standard Euler rotation of first
$\phi$ about the $z$-axis, then $\theta$ about the rotated $y$-axis,
with $\phi$ and $\theta$ the azimuthal and polar position of the observer.
Expressed algebraically:
\begin{align}
  \label{eqn:rotation_connection}
  \Omega_x &= \Omega_{x'}\cos\theta\cos\phi-\Omega_{y'}\sin\phi+\Omega_{z'}\sin\theta\cos\phi,\nonumber\\
  \Omega_y &= \Omega_{x'}\cos\theta\sin\phi+\Omega_{y'}\cos\phi+\Omega_{z'}\sin\theta\sin\phi,\nonumber\\
  \Omega_z &= \Omega_{x'}\sin\theta+\Omega_{z'}\cos\theta.
\end{align}

The scale factors $q$ and $s$ are functions of the neutrino
direction $\Omega_i$, the observer's velocity $u_\alpha$,
and the spacetime metric.
In the case of a fluid observer, specified by an arbitrary $W$ and $u_i$,
$q$ and $s$ are given by
\begin{align}
  \label{eqn:q_general}
  q
  &= W\alpha\Bigg(2\beta^i\Omega_i W^2(\beta^i\Omega_i-1) \nonumber\\
  &\qquad\qquad - 2\Omega_i u_j g^{ij} W \alpha(\beta^i\Omega_i-1)\nonumber\\
  &\qquad\qquad + \alpha^2\left((\Omega_i u_j g^{ij})^2+\Omega_i\Omega_j g^{ij}W^2\right)\Bigg)^{-1/2}\\
  \label{eqn:s_general}
  s
  &= \beta^i\Omega_i -\Omega_i u_j g^{ij}.
\end{align}
In the case of an Eulerian observer, $W=1$ and $u_i=0$,
and these expressions simplify considerably:
\begin{align}
  \label{eqn:q_eulerian}
  \tilde{q}
  &= \alpha\left(2\beta^i\Omega_i(\beta^i\Omega_i-1)+\alpha^2\Omega_i\Omega_j g^{ij}\right)^{-1/2}\\
  \label{eqn:s_eulerian}
  \tilde{s}
  &= \beta^i\Omega_i.
\end{align}
